Preparación de datos

### 2023-08-24 paper-blind.Rmd
entra JDiego

---
 title: "Subjective status, Social Class and meritocracy perception"
 # author: Julio Iturra ^[Bremen International Graduate School of Social Sciences (jiiturra@uni-bremen.de)]
 bibliography: "input/bib/ess-merit.bib"
 csl: "input/bib/apa6.csl"
 css: "input/css/custom.css"
 output:
     bookdown::html_document2:
          number_sections: false
     bookdown::pdf_document2:
         template: default
         toc: false
         number_sections: false
     bookdown::word_document2:
          toc: true
 geometry: margin=1in
 header-includes:
   # - \usepackage[spanish,es-tabla,es-nodecimaldot]{babel} # para decimales con commas en vez de punto
   - \usepackage{times}           # Times New Roman
   - \usepackage{caption}
   - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
   - \usepackage{graphicx}
   - \usepackage{float}
   - \usepackage{booktabs}
   - \usepackage{longtable}
   - \usepackage{array}
   - \usepackage{multirow}
   - \usepackage{wrapfig}
   - \usepackage{float}
   - \usepackage{colortbl}
   - \usepackage{pdflscape}
   - \usepackage{tabu}
   - \usepackage{threeparttable}
 linestretch: '1.5'
 fontsize: 12pt
 link-citations: yes
 linkcolor: blue
 editor_options:
   chunk_output_type: inline
 abstract: "Perceived meritocracy is understood as how societies describe the functioning of meritocracy, the ideal that promotes the distribution of resources in society based on individual effort and ability. Said this, are subjective and objective positions associated with perceived meritocracy? To what extent is objective position's influence on meritocracy perception moderated by subjective status? To answer these questions, we use nationally representative data from the 2018 wave of the Longitudinal Social Survey of Chile (N = 2,652). Regression analysis show that subjective social status has a robust positive association with perceived meritocracy, in line with the hypothesis. However, both positive and negative associations of income and occupational social class on perceived meritocracy were not confirmed. However, the hypothesis for education was confirmed, where higher educational attainment is associated with lower perceived meritocracy. Finally, we found that subjective status increases perceived meritocracy for those with higher incomes, tertiary education, and higher occupational social class."  
---

@@

The literature on subjective meritocracy has paid more attention to the structural position of individuals as an explanatory factor for perceptions of meritocracy. How individuals experience social inequality is affected by the status of the individuals' or groups [@castillo_perception_2022]. For example, perceived economic inequality in everyday life diminishes justification of inequality and increases demand for redistribution [@garcia-castro_perceiving_2020]. Nevertheless, it has been demonstrated that meritocracy primes reinforce negative evaluations of economically disadvantaged groups and strengthens internal attribution [@madeira_primes_2019].

@@

```{r tmodel, fig.cap='Theoretical model',out.height='40%', fig.align='center'}
 knitr::include_graphics(path = "input/images/scheme.PNG")
```

# Data, variables and methods

@@

```{r plot-likert, fig.height=3, fig.width=10, fig.cap='Perceived meritocracy items'}
 load(here::here("input/data/proc/elsoc_w03.RData"))   
 pacman::p_load(dplyr,sjPlot,ggplot2)

  elsoc_w03<- elsoc_w03 %>%
    dplyr::select(meffort,mtalent,
           inc10h.imp,edcine2,egpisko05,ess,lngap_perc,
           sexo,edad,ppolcat) %>% 
        na.omit() 
 dat_likert_s1 <-
   dplyr::select(elsoc_w03,meffort,mtalent)
 dat_likert_s1 <- sjmisc::rec(dat_likert_s1, rec="rev", append=FALSE)

@@

```{r ols-merit, echo=FALSE, results='asis'}
pacman::p_load(ordinal,texreg,dplyr) 
load(here::here("input/data/proc/elsoc_w03.RData"))
elsoc_w03$merit <- (as.numeric(elsoc_w03$meffort)+as.numeric(elsoc_w03$mtalent))/2 
  elsoc_w03<- elsoc_w03 %>%
    dplyr::select(merit,inc10h.imp,edcine2,egpisko05,ess,lngap_perc,
           sexo,edad,ppolcat) %>% 
        na.omit()

@@

>>>>

---
title: "Subjective status, Social Class and meritocracy perception"
linestretch: '1.5'          # interlineado
link-citations: yes         # citas linkeadas
fontsize: 12pt
linkcolor: blue
author:
- name: Julio Iturra ^[Corresponding author - jiturra@bigsss.uni-bremen.de]
  affiliation: Bremen International Graduate School of Social Sciences
- name: Juan Carlos Castillo
  affiliation: Universidad de Chile
bibliography: "input/bib/ess-merit.bib"
csl: "input/bib/apa6.csl"
css: "input/css/custom.css"
output:
  bookdown::word_document2:
    reference_docx: input/template.docx
    toc: false
    number_sections: false
  bookdown::html_document2:
    number_sections: false
  bookdown::pdf_document2:
    template: null
    toc: false
    keep_tex: false
    number_sections: false
    pandoc_args:
    - --template=input/mytemplate.tex #custom template para usar autores con afiliacion
geometry: margin=1in
header-includes:
  # - \usepackage[spanish,es-tabla,es-nodecimaldot]{babel} # para decimales con commas en vez de punto
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \hypersetup {colorlinks = true, linkcolor = blue, urlcolor = blue}  
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
lang: en-US
abstract: "Perceived meritocracy is understood as how societies describe the functioning of meritocracy, the ideal that promotes the distribution of resources in society based on individual effort and ability. Said this, are subjective and objective positions associated with perceived meritocracy? To what extent is objective position's influence on meritocracy perception moderated by subjective status? To answer these questions, we use nationally representative data from the 2018 wave of the Longitudinal Social Survey of Chile (N = 2,652). Regression analysis show that subjective social status has a robust positive association with perceived meritocracy, in line with the hypothesis. However, both positive and negative associations of income and occupational social class on perceived meritocracy were not confirmed. However, the hypothesis for education was confirmed, where higher educational attainment is associated with lower perceived meritocracy. Finally, we found that subjective status increases perceived meritocracy for those with higher incomes, tertiary education, and higher occupational social class."  
keywords: subjective social status, meritocracy perception, social class
classoption: a4paper
---

@@

The literature on subjective meritocracy has paid more attention to the structural position of individuals for explaning the emergence of meritocracy perceptions, assuming that structural positions and groups stratify [@castillo_perception_2022]. For example, perceived economic inequality in everyday life diminishes justification of inequality and increases demand for redistribution [@garcia-castro_perceiving_2020]. Nevertheless, it has been demonstrated that meritocracy primes reinforce negative evaluations of economically disadvantaged groups and strengthens internal attribution [@madeira_primes_2019].

@@

```{r tmodel, fig.cap='Theoretical model',out.height='30%', fig.align='center'}
 knitr::include_graphics(path = "input/images/scheme.PNG")
```

\pagebreak
# Data, variables and methods

@@

```{r plot-likert, fig.height=3, fig.width=10, fig.cap='Perceived meritocracy items'}
 load(here::here("input/data/proc/elsoc_w03.RData"))   
 pacman::p_load(dplyr,sjPlot,ggplot2)
 elsoc_w03$Q05h[elsoc_w03$Q05h=="No responde"] <- NA

  elsoc_w03<- elsoc_w03 %>%
    dplyr::select(meffort,mtalent,inc10h.imp,edcine2,class6,class3,ess,
                  egpisko05,Q05h,sexo,edad,ppolcat) %>% 
        na.omit() 
 dat_likert_s1 <-
   dplyr::select(elsoc_w03,meffort,mtalent)
 dat_likert_s1 <- sjmisc::rec(dat_likert_s1, rec="rev", append=FALSE)

@@

```{r ols-merit, echo=FALSE, results='asis'}
pacman::p_load(ordinal,texreg,dplyr,forcats,sjPlot) 
load(here::here("input/data/proc/elsoc_w03.RData"))
elsoc_w03$merit <- (as.numeric(elsoc_w03$meffort)+as.numeric(elsoc_w03$mtalent))/2 
elsoc_w03$class5 <- car::recode(
  elsoc_w03$class11d,
  as.factor = T,
  levels = c(
    'Service class (I - II)',
    'Routine non-manual (IIIa+b)',
    'Petty-bourgeoise (IVa+b+c)',
    'Skilled manual workers and supv. (V+VI)',
    'Non-skilled manual workers (VIIa+b)'
  ),
  "1:2='Service class (I - II)';
   3:4='Routine non-manual (IIIa+b)';
   5:7='Petty-bourgeoise (IVa+b+c)';
   8:9='Skilled manual workers and supv. (V+VI)';
   10:11='Non-skilled manual workers (VIIa+b)'"
)
elsoc_w03$Q05h[elsoc_w03$Q05h=="No responde"] <- NA
  elsoc_w03<- elsoc_w03 %>%
    dplyr::select(merit,meffort,mtalent,inc10h.imp,edcine2,class6,class3,ess,egpisko05,Q05h,class5,
           sexo,edad,ppolcat) %>% 
        na.omit()

table(elsoc_w03$class5,elsoc_w03$egpisko05)

elsoc_w03$ess_cwc <- mean(elsoc_w03$ess) - elsoc_w03$ess

elsoc_w03$ess5 <- car::recode(elsoc_w03$ess,"0:2=1;2:3=2;4:5=3;6:7=4;8:10=5")

table(elsoc_w03$class5)
elsoc_w03$class6 <- fct_rev(elsoc_w03$class6)  
elsoc_w03$class3 <- fct_rev(elsoc_w03$class3)  
elsoc_w03$class5 <- fct_rev(elsoc_w03$class5)  
table(elsoc_w03$Q05h)

@@

tab_model(lm(merit ~Q05h+edad+sexo,data= elsoc_w03),
              lm(merit ~edcine2+edad+sexo,data= elsoc_w03),
              lm(merit ~class3+edad+sexo,data= elsoc_w03),
              lm(merit ~Q05h+edcine2+class3+ess5+edad+sexo+ppolcat,data= elsoc_w03),
              lm(merit ~Q05h*ess5+edcine2+class3+edad+sexo+ppolcat,  data= elsoc_w03),  
              lm(merit ~Q05h+edcine2*ess5+class3+edad+sexo+ppolcat,  data= elsoc_w03),                
              lm(merit ~Q05h+edcine2+class3*ess5+edad+sexo+ppolcat,  data= elsoc_w03),
          show.ci = F,show.intercept = F,show.p = T,wrap.labels = 200,
          digits.p = 2,digits = 2,p.style = c("stars"),
          rm.terms = c("edad","sexoFemale"))

table(elsoc_w03$class3,elsoc_w03$ess5)

### 2023-08-24 production/prod_prep-data.Rmd
entra JDiego

>>>>

```{r egp-latam}
source(file = "production/prod_egp-respondent.R") #class scheme - respondent
source(file = "production/prod_egp-household.R")#class scheme - head household

#if NA in class of respondent, replace with household

sjmisc::frq(elsoc$clase11e)
sjmisc::frq(elsoc$clase11h)
elsoc$class11 <- ifelse(is.na(elsoc$clase11e),elsoc$clase11h,elsoc$clase11e)
sjmisc::frq(elsoc$class11)

# if NA in respondents it is assumed to be lower than head of household
# know, with the cases in we have information for respondent and head of household
# we replace the value by the highest one, just for those

# Replace the highest class by dominance criteria 
# Keep the highest of two values:
elsoc$class11d <- pmax(elsoc$class11, #social class respondent + household vs:
                       elsoc$clase11h,#social class household
                       na.rm = T) # if NA, do nothing

elsoc$class11d <- sjlabelled::set_labels(elsoc$class11d,labels = sjlabelled::get_labels(elsoc$clase11e)) 
sjmisc::frq(elsoc$class11d)
elsoc_w03$class11d <- elsoc$class11d

# Class Scheme 3 classes
elsoc_w03$class3 <- car::recode(elsoc_w03$class11d ,as.factor = T,
                            levels =c('Service Class (I+II)',
                                      'Intermediate class (III+IV)',
                                      'Working Class (V+VI+VII)'), 
                            "1:2='Service Class (I+II)';
                             3:7='Intermediate class (III+IV)';
                             8:11='Working Class (V+VI+VII)'")
sjmisc::frq(elsoc_w03$class3)

# Class Scheme 6 classes
elsoc_w03$class6 <- car::recode(
  elsoc_w03$class11d,
  as.factor = T,
  levels = c(
    'Service class (I - higher grade)',
    'Service class (II - lower grade)',
    'Routine non-manual (IIIa+b)',
    'Petty-bourgeoise (IVa+b+c)',
    'Skilled manual workers and supv. (V+VI)',
    'Non-skilled manual workers (VIIa+b)'
  ),
  "1='Service class (I - higher grade)';
   2='Service class (II - lower grade)';
   3:4='Routine non-manual (IIIa+b)';
   5:7='Petty-bourgeoise (IVa+b+c)';
   8:9='Skilled manual workers and supv. (V+VI)';
   10:11='Non-skilled manual workers (VIIa+b)'"
)

# Class Scheme 6 classes
elsoc_w03$class5 <- car::recode(
  elsoc_w03$class11d,
  as.factor = T,
  levels = c(
    'Service class (I - II)',
    'Routine non-manual (IIIa+b)',
    'Petty-bourgeoise (IVa+b+c)',
    'Skilled manual workers and supv. (V+VI)',
    'Non-skilled manual workers (VIIa+b)'
  ),
  "1:2='Service class (I - II)';
   3:4='Routine non-manual (IIIa+b)';
   5:7='Petty-bourgeoise (IVa+b+c)';
   8:9='Skilled manual workers and supv. (V+VI)';
   10:11='Non-skilled manual workers (VIIa+b)'"
)

sjmisc::frq(elsoc_w03$class6)
```
